---
title: "San Francisco Crimes (impacted by COVID-19)"
author: "Max Woolf (@minimaxir)"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: yes
    theme: spacelab
    toc: True
  keep_md: True
---

This notebook is licensed under the MIT License. If you use the code or data visualization designs contained within this notebook, it would be greatly appreciated if proper attribution is given back to this notebook and/or myself. Thanks! :)

## Setup

```{r}
library(tidyverse)
library(scales)
library(lubridate)

sessionInfo()
```

```{r}
theme_set(theme_minimal(base_size=8, base_family="Source Sans Pro") +
            theme(plot.title = element_text(family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.1, unit='cm'), size=10),
                  axis.title.x = element_text(),
                  axis.title.y = element_text(),
                  axis.text.y = element_text(family="Fira Code", size=4),
                  axis.text.x = element_text(family="Fira Code", size=4),
                  plot.subtitle = element_text(family="Source Sans Pro Semibold", color="#969696", size=8),
                  plot.caption = element_text(color="#969696"),
                  legend.title = element_text(),
                  legend.key.width = unit(0.25, unit='cm')))
```


# Load Data

```{r}
file_path <- "/Users/maxwoolf/Downloads/Police_Department_Incident_Reports__2018_to_Present.csv"
start_date <- "2020-03-01"
end_date <- "2020-03-28"   # last full day of data

df <- read_csv(file_path) %>%
  filter(`Incident Date` >= start_date, `Incident Date` <= end_date)

df %>% head()
```

Filter to only categories w/ atleast 10 at one day to clean up the cart.


```{r}
df_agg <- df %>%
  group_by(date = `Incident Date`, category = `Incident Category`) %>%
  tally() %>%
  arrange(category, date)

df_agg %>% head()
```

Filter to only categories w/ atleast 10 at one day to clean up the cart.

```{r}
categories <- df_agg %>%
  filter(n >= 10) %>%
  pull(category) %>%
  unique() %>%
  filter()

categories
```

Remove `Offences Against The Family And Children` because it's too verbose, and removing it gets the chart to a nice 4x4 square.


```{r}
sip_date <- "2020-03-16"

plot <-
  ggplot(
    df_agg %>% filter(
      category %in% categories,
      category != "Offences Against The Family And Children"
    ),
    aes(x = date, y = n, color = category)
  ) +
  geom_vline(
    aes(xintercept = as.Date(sip_date)),
    color = "black",
    linetype = "dashed",
    size = 0.25
  ) +
  geom_line() +
  scale_x_date(
    date_breaks = '1 week',
    date_labels = "%b\n%d",
    minor_breaks = NULL
  ) +
  scale_y_continuous(labels = comma_format(accuracy = 1), minor_breaks = NULL) +
  scale_color_hue(l = 50, guide = F) +
  facet_wrap( ~ category, scales = "free") +
  labs(title = "Reported Daily Crime Incidents in San Francisco Before/After COVID-19 Shelter-In-Place",
       subtitle = "For March 2020. Vertical dotted line indicates SIP announcement on March 16th. Data via datasf.org.",
       x = "Date",
       y = "# Reported Incidents",
       caption = "Max Woolf — @minimaxir")

ggsave('sf-crimes-covid.png', plot, width = 6, height = 4)
```



# LICENSE

MIT License

Copyright (c) 2020 Max Woolf

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.